### 2017-04-15 paper/oopsla17-submission/discussion.tex
even more pruning

\mysection{Limitations and Threats to Validity}
\label{sec:discussion}

@@

\mysubsection{Limitations}
\label{sec:limitations}

@@

\mysubsection{Threats to Validity}
\label{sec:validity}

Although our experiments demonstrate that our technique can pinpoint type
errors more accurately than the state of the art and that our features are
relevant to blame assignment, our results may not generalize.

One threat to validity associated with supervised machine learning is
overfitting (\ie learning a model that is too complex with respect to
the data).
%
A similar issue that arises in machine learning is model stability (\ie
can small changes to the training set produce large changes in the model?).
%
We mitigate these threats by:
%
(1) using separate training and testing datasets drawn from distinct
student populations (\autoref{sec:quantitative}), demonstrating the
generality of our models; and
%
(2) via cross-validation on the joint dataset
(\autoref{sec:feature-utility}), which demonstrates the stability of our
models by averaging the accuracy of 10 models trained on distinct
subsets of the data.

Our benchmarks were drawn from students in an undergraduate course and
may not be representative of other student populations.
%
We mitigate this threat by including the largest empirical evaluation of
type error localization that we are aware of: over 4,500 pairs of
ill-typed programs and fixes from two instances of the course, with
programs from 102 different students.
%
We acknowledge, of course, that students are not industrial programmers
and our results may not translate to large-scale software development;
however, we are particularly interested in aiding novice programmers
as they learn to work inside the type system.

A related threat to construct validity is our definition of the immedate
next well-typed program as the intended ground truth answer (see
\autoref{sec:overview}, Challenge 2). Students may, in theory, submit
intermediate well-typed program ``rewrites'' between the original ill-typed
program and the final intended answer. Our approach to discarding outliers
(see \autoref{sec:evaluation}) is designed to mitigate this threat.

Our removal of program pairs that changed too much, where our oracle
could not identify the blame of the other tools, or where the other
tools timed out or encountered unsupported language features is another
threat to validity.
%
It is possible that including the programs that changed excessively
would hurt our models, or that the other tools would perform
better on the programs with unsupported language features.
%
We note however that
%
(1) outlier removal is a standard technique in machine learning
\ES{CITE?}; and
%
(2) our Top-1 accuracy margin is large enough that even if we assumed
that \sherrloc were perfect on all of the excluded programs, it would
only tie our \hiddenFH in Top-1 accuracy.
%
\ES{I think this is accurate, but should double check..}

Examining programs written in \ocaml as opposed to \haskell or any other
typed functional language poses yet another threat, common type errors
may differ in different languages.
%
\ocaml is, however, a standard target for research in type error
localization and thus our choice admits a direct comparison with prior
work.
%
Furthermore, the functional core of \ocaml that we support does not
differ significantly from the functional core of \haskell or SML, all of
which are effectively lambda calculi with a Hindley-Milner-style type
system.\footnote{\haskell's type classes are a notable exception, they
  are also known to cause confusing type errors and would be interesting
  to study as well.}

Finally, our use of student fixes as oracles for the source of type
errors assumes that, on average, students are able to correctly identify
the source of an error.
%
As the students are in the process of learning the language and type
system, this assumption may be faulty.
%
It may be that \emph{expert} users would disagree with many of the
student fixes, and that it is harder to learn a model of expert fixes,
or that the state of the art would be better at predicting expert fixes.
%
As we have noted before, we believe it is reasonable to use student
fixes as oracles because the student is the best judge of what she
\emph{meant} to do.

>>>>

\mysection{Limitations}
\label{sec:discussion}

@@

% \mysubsection{Limitations}
% \label{sec:limitations}

@@

### 2017-04-15 paper/oopsla17-submission/evaluation.tex
even more pruning

\mypara{Feature Selection}
We extract a set of 277 features from each sub-expression in a student
program, including:
%
\begin{enumerate}
\item 44 local syntactic features. In addition to the syntax of \lang,
  we support the full range of arithmetic operators (integer and
  floating point), equality and comparison operators, character and
  string literals, and a user-defined |expr| type of simple arithmetic
  expressions. We discuss the challenge of supporting other
  user-defined types in \autoref{sec:discussion}.
\item 176 contextual syntactic features. For each sub-expression we
  additionally extract the local syntactic features of its parent and
  first, second, and third (left-to-right) children. If an expression
  does not have a parent or children, these features will simply be
  disabled. If an expression has more than three children, the
  classifiers will receive no information about the additional
  children.
\item 55 typing features. In addition to the types of \lang, we support
  |int|s, |float|s, |char|s, |string|s, and the user-defined |expr|
  mentioned above. These features are extracted for each sub-expression
  and for the contextual sub-expressions.
\item One feature denoting the size of each sub-expression.
\item One feature denoting whether each sub-expression is part of the
  minimal type error slice. We use this feature as a ``hard''
  constraint, sub-expressions that are not part of the minimal slice
  will be preemptively discarded. We justify this decision in
  \autoref{sec:feature-utility}.
\end{enumerate}

>>>>

\mypara{Feature Selection}
We extract a set of 277 features from each sub-expression in a student
program, including:
%
\begin{enumerate}
\item 44 local syntactic features. In addition to the syntax of \lang,
  we support the full range of arithmetic operators (integer and
  floating point), equality and comparison operators, character and
  string literals, and a user-defined % |expr| type of simple
  arithmetic
  expressions. We discuss the challenge of supporting other
  % user-defined
  types in \autoref{sec:discussion}.
\item 176 contextual syntactic features. For each sub-expression we
  additionally extract the local syntactic features of its parent and
  first, second, and third (left-to-right) children. If an expression
  does not have a parent or children, these features will simply be
  disabled. If an expression has more than three children, the
  classifiers will receive no information about the additional
  children.
\item 55 typing features. In addition to the types of \lang, we support
  |int|s, |float|s, |char|s, |string|s, and the user-defined |expr|
  mentioned above. These features are extracted for each sub-expression
  and its context. % for the contextual sub-expressions.
\item One feature denoting the size of each sub-expression.
\item One feature denoting whether each sub-expression is part of the
  minimal type error slice. We use this feature as a ``hard''
  constraint, sub-expressions that are not part of the minimal slice
  will be preemptively discarded. We justify this decision in
  \autoref{sec:feature-utility}.
\end{enumerate}

### 2017-04-15 paper/oopsla17-submission/learning.tex
even more pruning

A $\Model$ is produced by the $\trainsym$ function, which performs
supervised learning on a training set of feature vectors $\V$ and
(boolean) labels $\B$.
%
Once trained, a $\Model$ can be $\evalsym$uated on a new input vector,
producing the probability $\Runit$ that the blame label should be
applied.
%
We describe multiple types of $\Model$ and their instantiations of
$\trainsym$ and $\evalsym$ in \autoref{sec:models}.

Of course, the $\Model$ expects feature vectors $\V$ and blame labels
$\B$, but we are given program pairs.
%
So our first step must be to define a suitable translation from program
pairs to feature vectors and labels, \ie we must define the
$\extractsym$ function in \autoref{fig:api}.
%
We accomplish this by modeling features as real-valued functions of
expressions, and extract a feature vector for each \emph{subexpression}
of the ill-typed program (\autoref{sec:features}).
%
Then we define the blame labels for the training set to be the
subexpressions that changed between the ill-typed program and its
subsequent fix, and model $\blamesym$ as a function from a program pair
to the set of expressions that changed (\autoref{sec:labels}).
%
The $\extractsym$ function, then, extracts $\featuresym$ from each
subexpression and computes the blamed expressions according to
$\labelsym$.

>>>>

A $\Model$ is produced by the $\trainsym$ function, which performs
supervised learning on a training set of feature vectors $\V$ and
%(boolean)
labels $\B$.
%
Once trained, we can $\evalsym$uate a $\Model$ on a new input vector,
producing the confidence $\Runit$ that the blame label should apply. % be
%applied.
%
We describe multiple $\Model$s and their instantiations of
$\trainsym$ and $\evalsym$ (\autoref{sec:models}).

Of course, the $\Model$ expects feature vectors $\V$ and blame labels
$\B$, but we are given program pairs.
%
So our first step must be to define a suitable translation from program
pairs to feature vectors and labels, \ie we must define the
$\extractsym$ function in \autoref{fig:api}.
%
We model features as real-valued functions of
terms, and extract a feature vector for each \emph{subterm}
of the ill-typed program (\autoref{sec:features}).
%
Then we define the blame labels for the training set to be the
subexpressions that changed between the ill-typed program and its
subsequent fix, and model $\blamesym$ as a function from a program pair
to the set of expressions that changed (\autoref{sec:labels}).
%
The $\extractsym$ function, then, extracts $\featuresym$ from each
subexpression and computes the blamed expressions according to
$\labelsym$.

### 2017-04-15 paper/oopsla17-submission/related.tex
even more pruning

\mypara{Slicing}
Type error \emph{slicing}~\citep{Haack2003-vc,Tip2001-qp,Rahli2010-ps}
overcomes the traversal bias by extracting a complete and minimal subset
of program expressions that contribute to the type error, \ie all of the
expressions that are required for the error to manifest and no more.
%
Slicing techniques typically require rewriting the type checker with a
specialized constraint language and solver, though
\citet{Schilling2011-yf} shows how to turn any type checker into an
error slicer by treating it as a black-box.

While slicing techniques guarantee enough information to diagnose the
error, they can fall into the trap of providing \emph{too much}
information, producing a slice that is not significantly smaller than
the input program.

\mypara{Finding Likely Errors}
Thus, recent work has focused on finding the \emph{most likely} source
of a type error.
%
\citet{Zhang2014-lv} use Bayesian reasoning to search the constraint
graph for constraints that participate in many unsatisfiable paths and
relatively few satisfiable paths, based on the intuition that the
program should be mostly correct.
%
\citet{Pavlinovic2014-mr} translate the localization problem into a
MaxSMT problem, asking an off-the-shelf SMT solver to find the smallest
set of constraints that can be removed such that the resulting system is
satisfiable.
%
\citet{Loncaric2016-uk} improve the scalability of
\citeauthor{Pavlinovic2014-mr} by reusing the existing type checker as
an optimized theory solver in the Nelson-Oppen~\citeyear{Nelson1979-td}
style, and thus require only a MaxSAT solver.
%
All three of these techniques support \emph{weighted} constraints to
incorporate knowledge about the relative frequency of different errors,
but only \citeauthor{Pavlinovic2014-mr} use the weights, setting them to
the size of the sub-expression that induced the constraint.
%
In contrast, our classifiers learn a set of heuristics for predicting
the source of type errors by observing a set of ill-typed programs and
their subsequent fixes, in a sense using \emph{only} the weights and no
constraint solver.
%
It may be profitable to combine both approaches, \ie learn a set of good
weights for one of the above techniques from our training data.

\mypara{Explaining Type Errors}
In this paper we have focused solely on the task of \emph{localizing} a
type error, but a good error report should also \emph{explain} the
error.
%
\citet{Wand1986-nw}, \citet{Beaven1993-hb}, and \citet{Duggan1996-by}
attempt to explain type errors by collecting the chain of inferences
made by the type checker --- essentially the typing derivation --- and
presenting them to the user.
%
Such explanations can be lengthy, as an attempt to compress the
explanation \citet{Yang2000-kz} presents a visualization of the
inference process.
%
\citet{Gast2004-zd} produces a type error slice enhanced by arrows
showing the flow of data from sources with incompatible types to a
shared sink, borrowing the insight of dataflows-as-explanations from
\textsc{MrSpidey}~\citep{Flanagan1996-bu}.

@@

\mypara{Programmatic Explanations}
%
The best explanation of a type error, however, might be given by an
expert, \eg the compiler or library author.
%
\citet{Hage2006-hc} catalog a set of heuristics for
improving the quality of error messages by examining the errors made by
novices.
%
\citet{Heeren2003-db}, \citet{Christiansen2014-qc}, and
\citet{Serrano2016-oo} extend the ability to customize error messages to
library authors, enabling \emph{domain-specific} type errors.
%
The 8.0 release of the
Glasgow Haskell Compiler\footnote{\url{https://ghc.haskell.org/trac/ghc/wiki/Proposal/CustomTypeErrors}}
incorporates some of these ideas, allowing library authors to supply
custom errors when type-class resolution or type-family reduction fail,
but not for ordinary unification failures.

\mypara{Fixing Type Errors}
Some techniques go beyond explaining a type error or locating its
source, and actually attempt to \emph{fix} the error automatically.
%
\citet{Lerner2007-dt} searches for fixes to type errors by enumerating a
set of local mutations to the program and querying the type checker to
see if the error remains.
%
\citet{Chen2014-gd} use a notion of \emph{variation-based} typing to
track choices made by the type checker and enumerate potential type (and
expression) changes that would fix the error.
%
They also extend the algorithmic debugging technique of
\citeauthor{Chitil2001-td} by allowing the user to enter the expected
type of specific sub-expressions and suggesting fixes based on these
desired types \citeyear{Chen2014-vm}.
%
Our classifiers do not attempt to suggest fixes to type errors, but it
may be possible to do so by training a classifier to predict the
syntactic class of each expression in the \emph{fixed} program --- we
believe this is an exciting direction for future work.

\mysubsection{Fault Localization}
Given a defect, \emph{fault localization} is the task of identifying
``suspicious'' program elements (\eg lines, statements) that are likely
implicated in the defect (\ie that should be changed to fix the defect)
%
--- thus, type error localization can be viewed as an instance of fault
localization.
%
The best-known fault localization technique is likely Tarantula, which
uses a simple mathematical formula based on measured information from
dynamic normal and buggy runs~\citep{Jones2002-us}.
%
Other similar approaches, including those of \citet{Chen2002-qz} and
\citet{Abreu2006-fn,Abreu2007-mu} consider alternate features of
information or refined formulae and generally obtain more precise
results; see \citet{Wong2009-pd} for a survey.
%
While some researchers have approached such fault localization with an
eye toward optimality (\eg \citet{Yoo2013-rw} determine optimal
coefficients), in general such fault localization approaches are limited
by their reliance on either running tests or including relevant
features.
%
For example, Tarantula-based techniques require a normal and a buggy run
of the program.
%
By contrast, we consider incomplete programs with type errors that may
not be executed in any standard sense.
%
Similarly, the features available influence the classes of defects that
can be localized.
%
For example, a fault localization scheme based purely on control flow features
will have difficulty with cross-site scripting or SQL code injection
attacks, which follow the same control flow path on normal and buggy
runs (differing only in the user-supplied data).
%
Our feature set is comprised entirely of syntactic and typing features,
a natural choice for modeling type errors, but it would likely not
generalize to other types of defects.

>>>>

\mypara{Slicing}
Type error \emph{slicing}~\citep{Haack2003-vc,Tip2001-qp,Rahli2010-ps}
overcomes this bias by extracting a complete and minimal subset
of terms that contribute to the error, \ie all of the
terms that are required for it to manifest and no more.
%
Slicing typically requires rewriting the type checker with a
specialized constraint language and solver, though
\citet{Schilling2011-yf} shows how to turn any type checker into a
slicer by treating it as a black-box.
%
While slicing techniques guarantee enough information to diagnose the
error, they can fall into the trap of providing \emph{too much}
information, producing a slice that is not much smaller than
the input. %original. % input program.

\mypara{Finding Likely Errors}
Thus, recent work has focused on finding the \emph{most likely} source
of a type error.
%
\citet{Zhang2014-lv} use Bayesian reasoning to search the constraint
graph for constraints that participate in many unsatisfiable paths and
relatively few satisfiable paths, based on the intuition that the
program should be mostly correct.
%
\citet{Pavlinovic2014-mr} translate the localization problem into a
MaxSMT problem, asking an off-the-shelf solver to find the smallest
set of constraints that can be removed such that the resulting system is
satisfiable.
%
\citet{Loncaric2016-uk} improve the scalability of
\citeauthor{Pavlinovic2014-mr} by reusing the existing type checker as
a theory solver in the Nelson-Oppen~\citeyear{Nelson1979-td}
style, and thus require only a MaxSAT solver.
%
All three of these techniques support \emph{weighted} constraints to
incorporate knowledge about the frequency of different errors,
but only \citeauthor{Pavlinovic2014-mr} use the weights, setting them to
the size of the term that induced the constraint.
%
In contrast, our classifiers learn a set of heuristics for predicting
the source of type errors by observing a set of ill-typed programs and
their subsequent fixes, in a sense using \emph{only} the weights and no
constraint solver.
%
It may be profitable to combine both approaches, \ie learn a set of good
weights for one of the above techniques from our training data.

\mypara{Explaining Type Errors}
In this paper we have focused solely on the task of \emph{localizing} a
type error, but a good error report should also \emph{explain} the
error.
%
\citet{Wand1986-nw}, \citet{Beaven1993-hb}, and \citet{Duggan1996-by}
attempt to explain type errors by collecting the chain of inferences
made by the type checker %--- essentially the typing derivation ---
and presenting them to the user.
%
% Such explanations can be lengthy, as an attempt to compress the
% explanation \citet{Yang2000-kz} presents a visualization of the
% inference process.
%
\citet{Gast2004-zd} produces a slice enhanced by arrows
showing the dataflow from sources with different types to a
shared sink, borrowing the insight of dataflows-as-explanations from
\textsc{MrSpidey}~\citep{Flanagan1996-bu}.

@@

\mypara{Programmatic Explanations}
%
The best explanation of a type error, however, might be given by an
expert, \eg the compiler or library author.
%
\citet{Hage2006-hc} catalog a set of heuristics for
improving the quality of error messages by examining errors made by
novices.
%
\citet{Heeren2003-db}, \citet{Christiansen2014-qc}, and
\citet{Serrano2016-oo} extend the ability to customize error messages to
library authors, enabling \emph{domain-specific} errors.
%
The 8.0 release of the
Glasgow Haskell Compiler\footnote{\url{https://ghc.haskell.org/trac/ghc/wiki/Proposal/CustomTypeErrors}}
incorporates these ideas, allowing library authors to supply
custom errors when type-class resolution or type-family reduction fail,
but not for ordinary unification failures.

\mypara{Fixing Type Errors}
Some techniques go beyond explaining or locating a type error,
and actually attempt to \emph{fix} the error automatically.
%
\citet{Lerner2007-dt} searches for fixes by enumerating a
set of local mutations to the program and querying the type checker to
see if the error remains.
%
\citet{Chen2014-gd} use a notion of \emph{variation-based} typing to
track choices made by the type checker and enumerate potential
% type (and expression)
changes that would fix the error.
%
They also extend the algorithmic debugging technique of
\citeauthor{Chitil2001-td} by allowing the user to enter the expected
type of specific sub-expressions and suggesting fixes based on these
desired types \citeyear{Chen2014-vm}.
%
Our classifiers do not attempt to suggest fixes to type errors, but it
may be possible to do so by training a classifier to predict the
syntactic class of each expression in the \emph{fixed} program --- we
believe this is an exciting direction for future work.

\mysubsection{Fault Localization}
Given a defect, \emph{fault localization} is the task of identifying
``suspicious'' program elements (\eg lines, statements) that are likely
implicated in the defect %(\ie that should be changed to fix the defect)
%
--- thus, type error localization can be viewed as an instance of fault
localization.
%
The best-known fault localization technique is likely Tarantula, which
uses a simple mathematical formula based on measured information from
dynamic normal and buggy runs~\citep{Jones2002-us}.
%
Other similar approaches, including those of \citet{Chen2002-qz} and
\citet{Abreu2006-fn,Abreu2007-mu} consider alternate features of
information or refined formulae and generally obtain more precise
results; see \citet{Wong2009-pd} for a survey.
%
While some researchers have approached such fault localization with an
eye toward optimality (\eg \citet{Yoo2013-rw} determine optimal
coefficients), in general such fault localization approaches are limited
by their reliance on either running tests or including relevant
features.
%
For example, Tarantula-based techniques require a normal and a buggy run
of the program.
%
By contrast, we consider incomplete programs with type errors that may
not be executed in any standard sense.
%
Similarly, the features available influence the classes of defects that
can be localized.
%
For example, a fault localization scheme based purely on control flow features
will have difficulty with cross-site scripting or SQL code injection
attacks, which follow the same control flow path on normal and buggy
runs (differing only in the user-supplied data).
%
Our feature set is comprised entirely of syntactic and typing features,
a natural choice for modeling type errors, but it would likely not
generalize to other types of defects.
